A Stopped Negative Binomial Distribution

### 2016-05-30 stat-and-prob-letters/snb-paper.tex
More revisions.

\usepackage{lineno,hyperref}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{url}
\usepackage{subfig}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand*{\argminl}{\argmin\limits}
\newtheorem{prop}{Proposition}
\newtheorem{corollary}{Corollary}

@@

Consider a prototypical Phase II single-arm clinical trial in which 12 patients
are enrolled and treated. The binomial probability of success is $p=0.045$ for
each patient under the null hypothesis that treatment is no effective.
If two or more patients respond to the treatment then we reject this null
hypothesis, and the trial is deemed successful at significance level of $0.1$. 
If fewer than two respond then the null is not rejected.

If all 12 patients are enrolled at once, as in the classic
design, then the sample size is 12. However, as in most clinical trials, the
patients are enrolled sequentially, often with one patient's outcome realized
before the next one enters the trial. In the present example, observing two
successful patients allows us to reach one endpoint so the sample required
could be as small as two. Similarly 11
observed treatment failures also ends the trial. This sampling mechanism, in
which the experiment ends as soon as any of the endpoints is reached, is
call {\em curtailed sampling}. Under curtailed sampling the range of the
sample size for this trial is between two and 12.

@@

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients that
have been enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the third outcome and on in the tenth.

@@

In this work we derive the distribution of the number of enrollees needed
for either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB) and some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
The rest of this paper derives these results
and explores properties of the distribution.
The next section introduces our notation and basic results
including the density of the distribution along with a description of
its relation to other distributions. Section 2 derives the distribution
based on a defined Bernoulli process and gives some basic properties.
Section 3 provides a connection to the Binomial tail probability.
Section 4 derives the posterior distribution using a Beta prior distribution
for $p$. Section 5 provides a brief discussion on the use of the distribution
in clinical trials along with future avenues for generalization.

@@

The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
The probability mass function is
\begin{equation} \label{eqn:pmf}
\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability mass.

To prove (\ref{eqn:pmf}), consider the
process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \ I_{\{ k-t < X_k < s\}}.
\end{equation*}
At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz} 
we consider a graphical illustration of the plot $X_k$ against
$k$. If the outcome of the $k$th patient is a success then the process 
advances one diagonally in the
positive horizontal and vertical direction. If the $k$th patient fails
then the sample path advances in the positive horizontal direction only. The
process continues until either $X_k = s$ or $X_k = k-t$.

@@

To show that Equation \ref{eqn:pmf} sums to one, define
\begin{align} \label{eqn:sum_proof}
R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{align}
where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.

Then $R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
\citep{Olver2010} and recall this function satisfies 
$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Abramowitz1964}.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
This completes the proof (\ref{eqn:pmf}) is is the distribution of the
stopping time and a valid probability mass
function.
\end{proof}

@@

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
Similar derivations of the closely-related maximum negative binomial 
discrete distributions also appear in~\cite{Zhang2000}
and \cite{Zelterman2005}.
The maximum negative binomial is the smallest number of outcomes necessary to 
observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
number of outcomes to observe {\em either} $s$ successes or $t$ failures.

@@

There is a close connection between the tail areas of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to at most $s-1$ successes in a Binomial distribution. That is, the
probability that the number of successes is at least $s$ in the Binomial
model is the same that the trial succeeds (reaches $s$) in the SNB model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{align*}
\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \mathcal{I}_{1-p}.
\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}=s] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [Y \leq n\ |\ \text{\#success} = s] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

To illustrate this result, let us return to our initial example, shown
in Fig.~\ref{fig:snb_bin_compare}, 
where $s=2$, $t=12$, and $p=0.2$.  The probability masses represented in 
black are equal in a) and b) as are the masses in grey.
The probability that $s$
successes are reached in the SNB process is the same as the binomial 
probability of at least two successes. Likewise, the probability that $t$ 
failures are reached in the SNB process is the same as the binomial
probability of zero or one successes.

\begin{figure}[t!]
\centering
\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.2, 2, 12) with mass contributed from 
$s$ successes (black) or $t$ failures (grey) along with Bin(0.2, 13) with
at least 2 successes (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in Equation \ref{eqn:first_sum} satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.

A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

@@

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The posterior PMF of the Stopped Negative Binomial distribution with a 
Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
f(k | s, t, \alpha, \beta) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose k-t} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp.
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

The SNB is being used to generalize the Simon design by providing separate
criteria for futility and efficacy.
The Simon design tests for a single response probability and time to event
parameter for both the first
and second stage. The SNB is being used to generalize this design by 
allowing these parameters to vary for the two stages, and even allow
information about patients in the first stage to be used in the second
stage. Early results show the design provides
both early testing for futility along with the ability
to make statistically sound decisions sooner than traditional trials.

>>>>

\usepackage{lineno}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
%\usepackage{url}
\usepackage{subfig}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand*{\argminl}{\argmin\limits}
\newtheorem{prop}{Proposition}
\newtheorem{corollary}{Corollary}

@@

Consider a prototypical Phase II single-arm clinical trial in which 12 patients
are enrolled and treated. The binomial probability of success is $p=0.045$ for
each patient under the null hypothesis that treatment is not effective.
If two or more patients respond to the treatment then we reject this null
hypothesis and the trial is deemed successful at significance level of $0.1$. 
If fewer than two respond then the null is not rejected.

If all 12 patients are enrolled at once, as in the classic
design, then the sample size is 12. However, in most clinical trials the
patients are enrolled sequentially, often with one patient's outcome realized
before the next one enters the trial. In the present example, observing two
successful patients allows us to reach one endpoint so the sample required
could be as small as two. Similarly 11
observed treatment failures also ends the trial. This sampling mechanism, in
which the experiment ends as soon as a predefined endpoints is reached, is
call {\em curtailed sampling}. Under curtailed sampling the range of the
sample size for this trial is between two and 12.

@@

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients that
have been enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the fourth outcome and on in the eleventh.

@@

In this work we derive the distribution of the number of enrollees needed
for either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB) and some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
The rest of this paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the Binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the posterior distribution using a Beta prior distribution
for $p$. Section 6 provides a brief discussion on the use of the distribution
in clinical trials along with future avenues for generalization.

@@

The distribution of $\,Y\,$ has support on integer values in the range
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}.
\end{equation*}
The probability mass function is
\begin{equation} \label{eqn:pmf}
\mathbb{P} [Y=k] = S(k, p, s) \ I_{\{s \leq k\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is the {\emph indicator function}, taking a value 
of one if $f$ is true and zero otherwise, and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability mass.

To prove (\ref{eqn:pmf}), consider the
process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \ I_{\{ k-t < X_k < s\}}.
\end{equation*}
At each step a patient's outcome is measured. In Fig.~\ref{fig:kane_viz} 
we consider a graphical illustration of the plot $X_k$ against
$k$. If the outcome of the $k$th patient is a success then the process 
advances diagonally in the positive horizontal and vertical direction. 
If the $k$th patient fails
then the sample path advances in the positive horizontal direction only. The
process continues until either $X_k = s$ or $X_k = k-t$.

@@

To show that Equation \ref{eqn:pmf} sums to one, define
\begin{align} \label{eqn:sum_proof}
R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{align}
where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
Then $R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
\citep{Olver2010} and recall this function satisfies 
$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
This completes the proof (\ref{eqn:pmf}) is is the distribution of the
stopping time and a valid probability mass
function.
\end{proof}

@@

The SNB also generalizes both the minimum (riff-shuffle) and maximum negative
binomial distributions up to a translation of the support.
For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the
riff-shuffle, or minimum negative binomial distribution~\citep{Uppuluri1970}.
Similar derivations of the closely-related maximum negative binomial 
discrete distributions also appear in~\cite{Zhang2000}
and \cite{Zelterman2005}.
The maximum negative binomial is the smallest number of outcomes necessary to 
observe at least $s$ successes {\em and} $s$ failures. The SNB is the 
number of outcomes to observe {\em either} $s$ successes or $t$ failures.

@@

There is a close connection between the tail probabilities of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to at most $s-1$ successes in a Binomial distribution. That is, the
probability that the number of successes is at least $s$ in the Binomial
model is the same that the trial succeeds (reaches $s$) in the SNB model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{align*}
\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \mathcal{I}_{1-p}.
\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

To illustrate this result, let us return to our initial example, shown
in Fig.~\ref{fig:snb_bin_compare}, 
where $s=2$, $t=12$, and $p=0.045$.  The probability masses represented in 
black are equal in a) and b) as are the masses in grey.
The probability that $s$
successes are reached in the SNB process is the same as the binomial 
probability of at least two successes. Likewise, the probability that $t$ 
failures are reached in the SNB process is the same as the binomial
probability of zero or one successes.

\begin{figure}[t!]
\centering
\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.045, 2, 12) with mass contributed from 
$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 13) with
at least 2 successes (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in Equation \ref{eqn:first_sum} satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

@@

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The posterior PMF of the Stopped Negative Binomial distribution with a 
Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
f(k | s, t, \alpha, \beta) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose k-t} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp.
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

%The SNB is being used to generalize the Simon design by providing separate
%criteria for futility and efficacy.
%The Simon design tests for a single response probability and time-to-event
%parameter for both the first
%and second stage. The SNB is being used to generalize this design by 
%allowing these parameters to vary for the two stages, and even allow
%information about patients in the first stage to be used in the second
%stage. Early results show the design provides
%both early testing for futility along with the ability
%to make statistically sound decisions sooner than traditional trials.
